---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# nestyr

"Nesting" a data frame describes the process of combining values to create a 
single list-column variable of data frames *within* a data frame. The `tidyr` 
package makes this process easy with the `nest()` function. See an example here:
"http://r4ds.had.co.nz/many-models.html". 

Nesting in `tidyr` relies on grouping functions from the `dplyr` package to 
index the rows before nesting. However, as of September of 2018, `dplyr` is 
still unable to group by list columns (see 
https://github.com/tidyverse/tidyr/issues/249). In other words, you cannot nest
variables with a list column not included in your nest. This also means you 
cannot create multiple nests, as they are list columns.  

The `nestyr` package gets around this problem with the `nestyr::nest2()` 
function. Instead of grouping by list columns directly, it finds an MD5 hash for
all objects within any list columns before finding the group indices. 

The `nestyr::nest_cols()` function gets around the need to include a row id to 
nesting columns together while preserving rows. In other words, you can nest 
horizontally without thinking about row grouping. Like `nest2`, `nest_cols` also 
works with pre-existing list columns.  

## Example

Suppose we take the example provided by the "R for Data Science" book by Garret 
Germund and Hadley Wickham see (https://github.com/tidyverse/tidyr/issues/249). 
In this instance, however, let's say we wanted to nest together the columns 
`country` and `continent` together, and then nest together the remaining columns
by unique country-continent.

```{r example, message = F}
library(gapminder)
library(tidyverse)
library(nestyr)
gapminder %>% 
  nest_cols(country, continent, .key = "country_data") %>% 
  nest2(-country_data)
```

## Install

```{r eval = F}
library(devtools)
install_github("JakeNel28/nestyr")
```

## Acknowledgements

The package relies on much of the source code and some documentation used in 
`tidyr::nest()`, which was developed by Hadley Wickham "hadley@rstudio.com", 
Lionel Henry "lionel@rstudio.com", and RStudio https://www.rstudio.com/. I am 
very grateful to them for developing `tidyr`!
